---
title: "P8105 Homework 1"
author: "Shiyun Angel Cheng"
date: "2025-09-21"
output: github_document
---
## Problem 1 
```{r}
# packages installed in consel and load library
library(moderndive)
#access to data
data("early_january_weather")
```
Describe the dataset using inline code
```{r}
#check for variable names in the dataset 
names(early_january_weather)
```
The `early_january_weather` dataset contains the following variables: `r paste(names(early_january_weather), collapse = ", ")`. Some important variables include `temp` (temperature in Fahrenheit), `humid` (humidity level), and `time_hour` (timestamp).

```{r}
nrow(early_january_weather) #number of rows
ncol(early_january_weather) #number of columns 
mean(early_january_weather$temp, na.rm = TRUE) #average temperature
```
The dataset contains `r nrow(early_january_weather)` observations and `r ncol(early_january_weather)` variables. The average temperature is `r round(mean(early_january_weather$temp, na.rm = TRUE), 2)`.


Now we create a scatteplot.
```{r}
ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) +
  geom_point() +
  labs(title = "Temperature vs Time (colored by humidity)")
```

ggplot


